---
title: "Reproducible Research Week 2 Project"
author: "spoltone"
date: "28 09 2020"
output: html_document
---

## Introduction

This assignment makes use of data from personal activity monitoring device. This device collects data at 5 minute intervals each day.

The variables included in this dataset are:

steps: Number of steps taking in a 5-minute interval(missing values are coded as NA)
date: The date on which the measurement was taken in YYYY-MM-DD format
interval: Identifier for the 5-minute interval in which measurement was taken
The dataset is stored in a comma-separated-value (CSV) file and there are total of 17,568 observation in this dataset.

## Load and process the data

##### 1. Load the data and check

```{r echo=FALSE}

activityData <- read.csv('activity.csv')

head(activityData)

```

##### mean total number of steps taken per day


```{r}
stepsByDay <- tapply(activityData$steps, activityData$date, sum, na.rm=TRUE)

head(stepsByDay, 10)
```
##### 1. Histogram of the total number of steps taken each day

```{r}
library(ggplot2)

qplot(stepsByDay, xlab='Total steps per day', ylab='Frequency using binwith 500', binwidth=500)
```


##### Calculate and report the mean and median of the total number of steps taken per day

```{r}
stepsByDayMean <- mean(stepsByDay)
stepsByDayMedian <- median(stepsByDay)
```
* Mean: `r stepsByDayMean` 
* Median: `r stepsByDayMedian`

=====================================================

## Average daily activity pattern?

```{r}
averageStepsPerTimeBlock <- aggregate(x=list(meanSteps=activityData$steps), by=list(interval=activityData$interval), FUN=mean, na.rm=TRUE)
```

##### 1. Time series plot

```{r}
ggplot(data=averageStepsPerTimeBlock, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken") 
```

##### 2. The 5-minute interval on average across all the days in the dataset that contains the maximum number of steps?

```{r}
mostSteps <- which.max(averageStepsPerTimeBlock$meanSteps)
timeMostSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averageStepsPerTimeBlock[mostSteps,'interval'])
```

* Most Steps at: `r timeMostSteps`

=======================================================

## Missing values
##### 1. The total number of missing values in the dataset 

```{r}
numMissingValues <- length(which(is.na(activityData$steps)))
```

* Number of missing values: `r numMissingValues`

##### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
library(Hmisc)
activityDataImputed <- activityData
activityDataImputed$steps <- impute(activityData$steps, fun=mean)
```

##### 4. Histogram of the total number of steps taken each day 

```{r}
stepsByDayImputed <- tapply(activityDataImputed$steps, activityDataImputed$date, sum)
qplot(stepsByDayImputed, xlab='Total steps per day (Imputed)', ylab='Frequency using binwith 500', binwidth=500)
```

##### ... and Calculate and report the mean and median total number of steps taken per day. 

```{r}
stepsByDayMeanImputed <- mean(stepsByDayImputed)
stepsByDayMedianImputed <- median(stepsByDayImputed)
```
* Mean (Imputed): `r stepsByDayMeanImputed`
* Median (Imputed):  `r stepsByDayMedianImputed`

## Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
activityDataImputed$dateType <-  ifelse(as.POSIXlt(activityDataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

```{r}
averagedActivityDataImputed <- aggregate(steps ~ interval + dateType, data=activityDataImputed, mean)
ggplot(averagedActivityDataImputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```

